% ====================================================================
% A/L ICT - Lesson 6: Data Communication & Networking Cheat Sheet
% Optimized for A4 B&W Printing using LaTeX
% Designed for use on Overleaf.com
% ====================================================================

% --- Document Class & Packages ---
\documentclass[a4paper, 8pt]{extarticle}

% --- Geometry for Margins ---
\usepackage[a4paper, margin=1cm]{geometry}

% --- Essential Packages ---
\usepackage{multicol}        % For creating multi-column layouts
\usepackage[svgnames]{xcolor} % For colors (we'll use shades of gray)
\usepackage{tcolorbox}       % For creating styled boxes
\usepackage{enumitem}        % For customizing lists
\usepackage{tabularx}        % For better tables
\usepackage{graphicx}        % To scale images/symbols if needed

% --- Font Setup (for a clean, modern look) ---
\usepackage{fontspec}
\setmainfont{Inter}
\setmonofont{Roboto Mono}

% --- Custom Commands for Styling ---
\newcommand{\sectionheading}[1]{\large\textbf{#1}\par\noindent\rule{\linewidth}{0.4pt}}
\newcommand{\subsectionheading}[1]{\normalsize\textbf{#1}}
\tcbset{
  colback=gray!10, colframe=gray!75, boxrule=0.5pt,
  sharp corners, boxsep=4pt, left=4pt, right=4pt, top=4pt, bottom=4pt
}
\setlist[itemize]{leftmargin=*, noitemsep, topsep=2pt, label=\textbullet}

% --- Start of the Document ---
\begin{document}
\pagestyle{empty} % Removes page numbers

% --- Title ---
\begin{center}
    \fontsize{12pt}{14pt}\selectfont
    \textbf{Lesson 6: Data Communication \& Networking}
\end{center}
\vspace{1em}

% --- Main 2-Column Layout ---
\begin{multicols}{2}

% =============================================
% --- COLUMN 1 ---
% =============================================

\sectionheading{1. Physical Layer Concepts}
\vspace{0.5em}
\subsectionheading{Data Communication Model}
\begin{itemize}
    \item Sender $\rightarrow$ Message $\rightarrow$ Medium $\rightarrow$ Receiver $\rightarrow$ Protocol
\end{itemize}
\subsectionheading{Signal Properties}
\begin{itemize}
    \item \textbf{Types:} Analog (continuous) vs. Digital (discrete).
    \item \textbf{Properties:} Amplitude, Frequency, Wavelength, Phase.
    \item \textbf{Relationship:} \texttt{c = f * $\lambda$} (Speed = Freq * Wavelength)
\end{itemize}
\subsectionheading{Transmission Media}
\begin{itemize}
    \item \textbf{Guided:} Twisted Pair (UTP/STP), Coaxial, Fiber Optic.
    \item \textbf{Unguided:} Radio Waves, Microwaves.
\end{itemize}
\subsectionheading{Signal Impairment}
\begin{itemize}
    \item \textbf{Latency:} Delay in data travel.
    \item \textbf{Bandwidth:} Frequency range a medium can carry.
    \item \textbf{Noise:} Unwanted signal interference.
    \item \textbf{Attenuation:} Loss of signal strength over distance.
    \item \textbf{Distortion:} Change in the signal's shape.
\end{itemize}
\subsectionheading{Modulation \& Encoding}
\begin{itemize}
    \item \textbf{Modulation:} AM, FM, PM, PCM (Pulse Code Modulation).
    \item \textbf{Digital-to-Analog:} ASK, FSK, PSK.
    \item \textbf{Digital Encoding:} Manchester encoding (aids synchronization).
\end{itemize}

\vspace{1em}
\sectionheading{2. Data Link Layer Concepts}
\vspace{0.5em}
\subsectionheading{Network Topologies}
\begin{itemize}
    \item Point-to-Point, Bus, Star, Ring, Mesh.
\end{itemize}
\subsectionheading{LAN Hardware}
\begin{itemize}
    \item \textbf{NIC:} Connects device to network.
    \item \textbf{Hub:} Broadcasts data to all (Dumb).
    \item \textbf{Switch:} Sends data to intended port (Intelligent).
    \item \textbf{Bridge:} Connects two similar LANs.
\end{itemize}
\subsectionheading{Media Access Control (MAC)}
\begin{itemize}
    \item \textbf{MAC Address:} 48-bit unique hardware address (in 6 blocks).
    \item \textbf{Frame:} Data unit at Data Link layer, contains MAC of \textit{next adjacent node}.
    \item \textbf{Protocols:} Manage access to shared media (Evolution: ALOHA $\rightarrow$ Ethernet (CSMA/CD)).
    \item \textbf{Error Control:} Parity Check (detects single-bit errors).
\end{itemize}

\columnbreak % End Column 1, Start Column 2

% =============================================
% --- COLUMN 2 ---
% =============================================

\sectionheading{3. Network \& Transport Layers}
\vspace{0.5em}
\subsectionheading{IP Addressing}
\begin{itemize}
    \item \textbf{Versions:} IPv4 (32-bit) vs. IPv6 (128-bit).
    \item \textbf{IPv4 Classes:} A (1-126), B (128-191), C (192-223).
    \item \textbf{Private IP Ranges:} \texttt{10.0.0.0/8}, \texttt{172.16.0.0/12}, \texttt{192.168.0.0/16}.
    \item \textbf{Subnetting:} Dividing a large network into smaller sub-networks. Used with a \textbf{Subnet Mask}.
    \item \textbf{CIDR Notation:} e.g., /24.
\end{itemize}
\subsectionheading{Internetworking Hardware}
\begin{itemize}
    \item \textbf{Router:} Connects different networks, finds the best path.
    \item \textbf{Gateway:} Connects networks using different protocols.
\end{itemize}
\subsectionheading{Core Internet Concepts}
\begin{itemize}
    \item \textbf{Packet Switching:} Messages broken into small packets.
    \item \textbf{Best-Effort Delivery:} Network doesn't guarantee packet delivery.
\end{itemize}
\subsectionheading{Transport Layer Protocols}
\begin{itemize}
    \item \textbf{Ports \& Multiplexing:} How multiple apps use the network at once.
    \item \textbf{UDP (User Datagram Protocol):} Connectionless, no delivery guarantee. (Fast; for DNS, streaming).
    \item \textbf{TCP (Transmission Control Protocol):} Reliable, connection-oriented, ordered delivery. (Slower; for web, email).
\end{itemize}

\vspace{1em}
\sectionheading{4. Application, Security & Connectivity}
\vspace{0.5em}
\subsectionheading{Reference Models}
\begin{itemize}
    \item \textbf{OSI (7 Layers):} Conceptual model (Physical, Data Link, Network, Transport, Session, Presentation, Application).
    \item \textbf{TCP/IP (4 Layers):} Practical model used on the internet.
\end{itemize}
\subsectionheading{Application Layer Services}
\begin{itemize}
    \item \textbf{DHCP:} Automatically assigns IP addresses.
    \item \textbf{DNS:} Translates domain names (with TLDs like .com, .lk) to IP addresses.
    \item \textbf{NAT/Proxy:} Allows private IPs to access the internet.
\end{itemize}
\subsectionheading{Connectivity}
\begin{itemize}
    \item \textbf{ISP:} Provides internet access.
    \item \textbf{DSL/ADSL:} Broadband over phone lines. (Advantages: Independent service, more secure than cable).
    \item \textbf{Modem:} Converts signals for different media.
\end{itemize}
\subsectionheading{Network Security}
\begin{itemize}
    \item \textbf{Encryption:} Symmetric (one key) vs. Asymmetric (public/private keys).
    \item \textbf{Digital Signature:} Authenticates sender \& message integrity.
    \item \textbf{Threats:} Viruses, Trojans, Malware, Phishing.
    \item \textbf{Protection:} Firewalls, Antivirus software.
\end{itemize}

\end{multicols}
\end{document}
