% ====================================================================
% A/L ICT - Lesson 1: Core ICT Concepts Cheat Sheet
% Optimized for A4 B&W Printing using LaTeX
% Designed for use on Overleaf.com
% ====================================================================

% --- Document Class & Packages ---
\documentclass[a4paper, 8pt]{extarticle}

% --- Geometry for Margins ---
\usepackage[a4paper, margin=1cm]{geometry}

% --- Essential Packages ---
\usepackage{multicol}        % For creating multi-column layouts
\usepackage[svgnames]{xcolor} % For colors (we'll use shades of gray)
\usepackage{tcolorbox}       % For creating styled boxes
\usepackage{enumitem}        % For customizing lists

% --- Font Setup (for a clean, modern look) ---
\usepackage{fontspec}
\setmainfont{Inter}
\setmonofont{Roboto Mono}

% --- Custom Commands for Styling ---
\newcommand{\sectionheading}[1]{\large\textbf{#1}\par\noindent\rule{\linewidth}{0.4pt}}
\newcommand{\subsectionheading}[1]{\normalsize\textbf{#1}}
\tcbset{
  colback=gray!10, colframe=gray!75, boxrule=0.5pt,
  sharp corners, boxsep=4pt, left=4pt, right=4pt, top=4pt, bottom=4pt
}
\setlist[itemize]{leftmargin=*, noitemsep, topsep=3pt, label=\textbullet}

% --- Start of the Document ---
\begin{document}
\pagestyle{empty} % Removes page numbers

% --- Title ---
\begin{center}
    \fontsize{12pt}{14pt}\selectfont
    \textbf{Lesson 1: Core ICT Concepts}
\end{center}
\vspace{1em}

% --- Main 2-Column Layout ---
\begin{multicols}{2}

% =============================================
% --- COLUMN 1 ---
% =============================================

\sectionheading{1. Data, Information \& Technology}
\vspace{0.5em}
\subsectionheading{Data vs. Information}
\begin{itemize}
    \item \textbf{Data:} Raw facts. \textit{(Uncooked ingredients)}.
    \item \textbf{Information:} Data with meaning. \textit{(The finished cake)}.
    \item \textit{One system's Information can be another's Data.}
    \item \textbf{Valuable Info is:} Timely, Accurate, Complete, Relevant.
    \item \textbf{Big Data:} Google-scale data analysis.
\end{itemize}

\subsectionheading{The Need for Technology}
\begin{itemize}
    \item \textbf{Internet:} The global "network of networks".
        \begin{itemize}
            \item \textbf{TCP/IP:} Internet's Rulebook / Postal Service.
            \item \textbf{Services:} WWW, FTP, Email, Telnet, etc.
        \end{itemize}
    \item \textbf{Cloud Computing:} Using someone else's computer online.
        \begin{itemize}
            \item \textbf{SaaS:} Rent their finished software.
            \item \textbf{PaaS:} Rent their workshop.
            \item \textbf{IaaS:} Rent their land \& raw materials.
        \end{itemize}
\end{itemize}

\vspace{1em}
\sectionheading{2. The Computer System}
\vspace{0.5em}
\subsectionheading{Core Concepts}
\begin{itemize}
    \item \textbf{Abstract Model:} Input $\rightarrow$ Process $\rightarrow$ Output.
    \item \textbf{Firmware:} Software built into hardware. \textit{(e.g., TV remote's brain)}.
    \item \textbf{Liveware:} You, the human user.
\end{itemize}
\subsectionheading{Hardware (The physical parts)}
\begin{itemize}
    \item \textbf{Processing:} CPU (The main \textbf{Brain}), GPU (The \textbf{Graphics Brain}).
    \item \textbf{Memory:}
    \begin{itemize}
        \item \textbf{Cache:} CPU's ultra-fast \textbf{Notepad}.
        \item \textbf{RAM (Primary):} Temporary Workspace. \textbf{Volatile} (forgets when power is off).
        \item \textbf{Secondary:} Filing Cabinet. \textbf{Non-volatile} (remembers always).
    \end{itemize}
    \item \textbf{Networking Hardware:} NIC, Hub, Switch, Router, Gateway.
\end{itemize}
\subsectionheading{Software (The instructions)}
\begin{itemize}
    \item \textbf{System Software:} Manages the computer (OS, Utilities).
    \item \textbf{Application Software:} The "Apps" for tasks (Word, Chrome).
    \item \textbf{Licensing:}
    \begin{itemize}
        \item \textbf{Proprietary:} A "Secret Recipe".
        \item \textbf{FOSS (Open Source):} A "Public Recipe".
    \end{itemize}
\end{itemize}

\columnbreak % End Column 1, Start Column 2

% =============================================
% --- COLUMN 2 ---
% =============================================

\sectionheading{3. Data Processing In Action}
\vspace{0.5em}
\begin{tcolorbox}[title=\textbf{The Cycle}]
    Gathering $\rightarrow$ Validation $\rightarrow$ Processing $\rightarrow$ Output $\rightarrow$ Storage
\end{tcolorbox}
\subsectionheading{Gathering Methods}
\begin{itemize}
    \item \textbf{Manual:} Interviews, Questionnaires, Observation.
    \item \textbf{Automated:}
    \begin{itemize}
        \item \textbf{OMR:} The "MCQ Answer Sheet Reader".
        \item \textbf{OCR:} The "Image-to-Text Converter".
        \item \textbf{MICR:} The "Bank Cheque Reader".
    \end{itemize}
\end{itemize}
\subsectionheading{Validation (The "Data Quality Checkpoint")}
\begin{itemize}
    \item \textbf{Range Check:} Is the value within limits? (e.g., 0-100).
    \item \textbf{Presence Check:} Is it empty?
    \item \textbf{Type Check:} Is it a number or text?
\end{itemize}
\subsectionheading{Processing Modes}
\begin{itemize}
    \item \textbf{Batch:} "End-of-day" processing. Collect \& process in one go.
    \item \textbf{Real-time:} "Instant" processing. Immediate response.
\end{itemize}

\vspace{1em}
\sectionheading{4. ICT's Role \& Impact on Society}
\vspace{0.5em}
\subsectionheading{Key Applications}
\begin{itemize}
    \item \textbf{CAD (Computer-Aided Design):} The "Engineer's Digital Drawing Board".
    \item \textbf{RFID (Radio-Frequency Identification):} A "Wireless Barcode".
\end{itemize}
\subsectionheading{Key Societal Issues}
\begin{itemize}
    \item \textbf{Digital Divide:} The "Tech Haves vs. Tech Have-Nots".
    \item \textbf{E-waste:} Discarded electronics.
        \begin{itemize}
            \item \textbf{Green Computing:} "Eco-Friendly Tech".
        \end{itemize}
\end{itemize}
\subsectionheading{Ethical \& Legal Issues (The "Don'ts")}
\begin{itemize}
    \item \textbf{Piracy:} Stealing software/media.
    \item \textbf{Plagiarism:} Copying someone's work.
    \item \textbf{Phishing:} A "Fake Login" trap.
    \item \textbf{Copyright / Intellectual Property:} Creator's legal ownership.
    \item \textbf{Confidentiality:} Keeping data "Secret".
    \item \textbf{Netiquette:} Internet Ethics / Good Manners.
\end{itemize}

\end{multicols}
\end{document}
