% ====================================================================
% A/L ICT - Lesson 12: ICT for Business Cheat Sheet
% Optimized for A4 B&W Printing using LaTeX
% Based on the final revised plan.
% ====================================================================

% --- Document Class & Packages ---
\documentclass[a4paper, 8pt]{extarticle}

% --- Geometry for Margins ---
\usepackage[a4paper, margin=1cm]{geometry}

% --- Essential Packages ---
\usepackage{multicol}        % For creating multi-column layouts
\usepackage[svgnames]{xcolor} % For colors (we'll use shades of gray)
\usepackage{tcolorbox}       % For creating styled boxes
\usepackage{enumitem}        % For customizing lists
\usepackage{tabularx}        % For better tables

% --- Font Setup (for a clean, modern look) ---
\usepackage{fontspec}
\setmainfont{Inter}
\setmonofont{Roboto Mono}

% --- Custom Commands for Styling ---
\newcommand{\sectionheading}[1]{\large\textbf{#1}\par\noindent\rule{\linewidth}{0.4pt}}
\newcommand{\subsectionheading}[1]{\normalsize\textbf{#1}}
\tcbset{
  colback=gray!10, colframe=gray!75, boxrule=0.5pt,
  sharp corners, boxsep=4pt, left=4pt, right=4pt, top=4pt, bottom=4pt
}
\setlist[itemize]{leftmargin=*, noitemsep, topsep=3pt, label=\textbullet}

% --- Start of the Document ---
\begin{document}
\pagestyle{empty} % Removes page numbers

% --- Title ---
\begin{center}
    \fontsize{12pt}{14pt}\selectfont
    \textbf{Lesson 12: ICT for Business}
\end{center}
\vspace{1em}

% --- Main 2-Column Layout ---
\begin{multicols}{2}

% =============================================
% --- COLUMN 1 ---
% =============================================

\sectionheading{1. The Digital Marketplace \& Business Models}
\vspace{0.5em}
\subsectionheading{Digital Economy \& Business Transformation}
\begin{itemize}
    \item \textbf{Digital Economy:} An economy based on digital computing technologies, driven by the Internet.
    \item \textbf{Information Systems (IS):} The key ICT-based tool for business transformation.
    \item \textbf{Types of IS:}
    \begin{itemize}
        \item \textbf{TPS:} Transaction Processing Systems (e.g., a sale).
        \item \textbf{Decision Support:} \texttt{MIS}, \texttt{DSS}, \texttt{ESS}.
    \end{itemize}
\end{itemize}

\subsectionheading{Business Models in the Digital Economy}
\begin{itemize}
    \item \textbf{New Methods:} Reverse Auctions, Group Purchasing, e-Marketplace.
    \item \textbf{Organization Types:} Pure Brick, Brick and Click, Pure Click.
\end{itemize}

\subsectionheading{Role of ICT in Business Functions}
\begin{itemize}
    \item \textbf{Accounting}, \textbf{HR} (\texttt{HRIS}, self-service portals), \textbf{Production} (Industry 4.0), \textbf{Marketing}, \textbf{SCM} (Inter-organizational systems), \textbf{Communication} (Teleworking).
\end{itemize}

\subsectionheading{Secure Payment Mechanisms}
\begin{itemize}
    \item \textbf{Payment Gateway:} Securely handles credit card verification.
    \item \textbf{Third-party Systems:} e.g., PayPal (mediates payment).
    \item \textbf{Mechanisms:} Data Encryption.
    \item \textbf{Currencies:} Virtual and Crypto (e.g., Bitcoin).
\end{itemize}

\subsectionheading{Threats and Opportunities}
\begin{itemize}
    \item Privacy, Security, Product Commercialization.
\end{itemize}

\vspace{1em}
\sectionheading{3. E-Marketing \& Competitive Advantage}
\vspace{0.5em}
\subsectionheading{E-marketing}
\begin{itemize}
    \item Using the internet for marketing (promotions, PR, support).
    \item \textbf{Use of ICT:} Web Advertising, Data Mining.
\end{itemize}

\subsectionheading{Mobile Marketing}
\begin{itemize}
    \item A form of e-marketing via mobile devices.
    \item \textbf{Location-Based Services:}
    \begin{itemize}
        \item \textbf{Geosocial:} Find people with similar interests nearby.
        \item \textbf{Geoadvertising:} Target users in a specific location.
        \item \textbf{Geoinformation:} Provide info about a location.
    \end{itemize}
\end{itemize}

\subsectionheading{Database Marketing}
\begin{itemize}
    \item A direct marketing technique using customer databases.
    \item Uses data mining to create customer profiles for personalized offers.
\end{itemize}

\subsectionheading{Gaining Competitive Advantages through ICT}

\columnbreak % End Column 1, Start Column 2

% =============================================
% --- COLUMN 2 ---
% =============================================

\sectionheading{2. E-Commerce \& E-Business Operations}
\vspace{0.5em}
\subsectionheading{E-Commerce vs. E-Business}
\begin{itemize}
    \item \textbf{E-Business:} Broader concept. Includes all internal and external electronic business processes.
    \begin{itemize}
        \item \textbf{Architecture:} Often via \textbf{ERP} systems, integrating modules (Sales, Finance) with a central \textbf{Corporate Database}.
    \end{itemize}
    \item \textbf{E-Commerce:} A subset of E-Business, focusing on buying and selling online.
\end{itemize}

\subsectionheading{Types of E-Commerce Transactions}
\begin{tcolorbox}
\begin{itemize}
    \item \textbf{B2B:} Business-to-Business
    \item \textbf{B2C:} Business-to-Consumer
    \item \textbf{C2C:} Consumer-to-Consumer
    \item \textbf{C2B:} Consumer-to-Business
    \item \textbf{B2E:} Business-to-Employee
    \item \textbf{G2C:} Government-to-Citizen
\end{itemize}
\end{tcolorbox}

\subsectionheading{Internet and Business (E-Commerce Components)}
\begin{itemize}
    \item \textbf{Virtual Storefronts:} Online shop display.
    \item \textbf{Information Brokers:} Filter information for consumers.
    \item \textbf{Online Marketplace:} Platform for multiple sellers (e.g., eBay).
    \item \textbf{Content Provider:} Offers digital content (e.g., iTunes).
    \item \textbf{Online Service Provider:} Offers services (e.g., Google).
    \item \textbf{Portal:} Entry point to e-commerce (e.g., Yahoo).
    \item \textbf{Virtual Community:} Online social networks.
\end{itemize}

\subsectionheading{Advantages and Disadvantages of E-Commerce}
\begin{itemize}
    \item \textit{(Briefly list key pros and cons based on syllabus understanding)}
\end{itemize}

\subsectionheading{E-Commerce Trends}
\begin{itemize}
    \item Becoming more social, mobile-based (\texttt{m-commerce}), etc.
\end{itemize}


\end{multicols}
\end{document}
