% ====================================================================
% A/L ICT - Lesson 1: Core ICT Concepts Cheat Sheet (Improved Version)
% Optimized for A4 B&W Printing, Single Page
% Designed for use on Overleaf.com
% ====================================================================

% --- Document Class & Packages ---
\documentclass[a4paper, 8pt]{extarticle}

% --- Geometry for Margins (Crucial for fitting content) ---
\usepackage[a4paper, margin=1cm, top=1.2cm, bottom=1.2cm]{geometry}

% --- Essential Packages ---
\usepackage{multicol}        % For creating multi-column layouts
\usepackage[svgnames]{xcolor} % For colors (used for black/gray)
\usepackage{tcolorbox}       % For creating styled boxes
\usepackage{enumitem}        % For customizing lists
\usepackage{amssymb}         % For symbols like checkmark

% --- Font Setup (for a clean, modern look) ---
\usepackage{fontspec}
\setmainfont{Inter}
\setmonofont{Roboto Mono}

% --- Custom Commands for Styling ---
% Command for main section headings with a line underneath
\newcommand{\sectionheading}[1]{\large\textbf{#1}\par\noindent\rule{\linewidth}{0.4pt}}
% Command for subsection headings
\newcommand{\subsectionheading}[1]{\normalsize\textbf{#1}}

% --- tcolorbox configuration for B&W printing ---
% No background color (colback=white), only a thin black frame
\tcbset{
  colback=white, colframe=black, boxrule=0.5pt,
  sharp corners, boxsep=4pt, left=4pt, right=4pt, top=4pt, bottom=4pt
}

% --- List configuration for compactness ---
% No extra space between list items
\setlist[itemize]{leftmargin=*, noitemsep, topsep=2pt, label=\textbullet}
\setlist[enumerate]{leftmargin=*, noitemsep, topsep=2pt}


% --- Start of the Document ---
\begin{document}
\pagestyle{empty} % Removes page numbers

% --- Title ---
\begin{center}
    \fontsize{12pt}{14pt}\selectfont
    \textbf{A/L ICT | Lesson 1: ICT concepts}
\end{center}
\vspace{0.5em}

% --- Main 2-Column Layout ---
\begin{multicols}{2}

% =============================================
% --- COLUMN 1 ---
% =============================================

\sectionheading{1. Data, Information \& Technology}
\vspace{0.5em}
\subsectionheading{Data, Information, Knowledge}
\begin{itemize}
    \item \textbf{Data:} Raw, unorganized facts. \textit{(e.g., 19, 25, 22)}.
    \item \textbf{Information:} Processed, meaningful data. \textit{(e.g., "The average temperature is 22°C")}.
    \item \textit{Note: One system's Information can be another's Data.}
    \item \textbf{Data Types:}
        \begin{itemize}
            \item \textbf{Quantitative:} Numerical, measurable. \textit{(e.g., Marks)}.
            \item \textbf{Qualitative:} Descriptive, non-numerical. \textit{(e.g., Goodwill)}.
        \end{itemize}
    \item \textbf{Data Life Cycle:} Creation $\rightarrow$ Management $\rightarrow$ Removal.
\end{itemize}

\subsectionheading{Value of Information}
\begin{itemize}
    \item \textbf{Characteristics:} Accurate, Complete, Timely, Relevant, Understandable.
    \item \textbf{Golden Rule:} Value is highest at creation, decreases over time.
    \item \textbf{Big Data:} Extremely large data sets requiring advanced tools for analysis.
\end{itemize}

\subsectionheading{Need for Technology}
\begin{itemize}
    \item \textbf{Manual Processing Drawbacks:} Slow, error-prone, inconsistent data, poor sharing.
    \item \textbf{Internet:} Global "network of networks". Uses \texttt{TCP/IP} protocol.
    \begin{itemize}
        \item \textbf{Services:} WWW, Email, FTP, Telnet, VoIP, etc.
    \end{itemize}
    \item \textbf{Mobile Tech:}
        \begin{itemize}
            \item \textbf{Communication:} Wireless data/voice transmission.
            \item \textbf{Computing:} Using devices without a fixed physical link.
        \end{itemize}
    \item \textbf{Cloud Computing:} Online access to shared computing resources.
    \begin{itemize}
        \item \textbf{IaaS:} Rent infrastructure. \textit {(Land  \& Bricks)}.
        \item \textbf{PaaS:} Rent a platform. \textit{(Workshop)}.
        \item \textbf{SaaS:} Rent software. \textit{(Finished Product)}.
    \end{itemize}
\end{itemize}

\vspace{1em}
\sectionheading{2. The Computer System}
\vspace{0.5em}
\begin{tcolorbox}[title=\textbf{System Definition}]
An interrelated set of components working together to achieve a common objective. \textbf{Model:} Input $\rightarrow$ Process $\rightarrow$ Output.
\end{tcolorbox}
\subsectionheading{System Components}
\begin{itemize}
    \item \textbf{Hardware:} Physical parts. (Input, Output, CPU, Memory).
    \item \textbf{Software:} Instructions. (System, Application).
    \item \textbf{Firmware:} Software embedded in hardware. \textit{(e.g., BIOS)}.
    \item \textbf{Liveware:} The human user.
\end{itemize}
\subsectionheading{Hardware Details}
\begin{itemize}
    \item \textbf{Processing:} CPU (main brain), GPU (graphics brain).
    \item \textbf{Memory Hierarchy:}
        \begin{itemize}
            \item \textbf{Cache:} CPU's ultra-fast notepad.
            \item \textbf{RAM (Primary):} Temp workspace, \textbf{Volatile}.
            \item \textbf{Secondary:} Permanent storage, \textbf{Non-volatile}.
        \end{itemize}
    \item \textbf{Networking:} NIC, Switch, Router, Gateway, Bridge.
\end{itemize}
\subsectionheading{Software Details}
\begin{itemize}
    \item \textbf{System Software:} Manages the computer.
        \begin{itemize}
            \item \textbf{Operating System (OS):} e.g., Windows, Linux.
            \item \textbf{Utilities:} e.g., Antivirus, Disk Tools.
            \item \textbf{Language Translators:} Compiler, Interpreter.
        \end{itemize}
    \item \textbf{Application Software:} User-task oriented. \textit{(e.g., Word)}.
    \item \textbf{Software Licensing:}
        \begin{itemize}
            \item \textbf{Proprietary:} Secret recipe. Limited rights.
            \item \textbf{FOSS (Open Source):} Public recipe. Free to use, modify, share.
        \end{itemize}
\end{itemize}

\columnbreak % End Column 1, Start Column 2

% =============================================
% --- COLUMN 2 ---
% =============================================

\sectionheading{3. Data Processing In Action}
\vspace{0.5em}
\begin{tcolorbox}[title=\textbf{Data Processing Cycle}]
Gathering $\rightarrow$ Validation $\rightarrow$ Processing $\rightarrow$ Output $\rightarrow$ Storage
\end{tcolorbox}
\subsectionheading{Gathering \& Input}
\begin{itemize}
    \item \textbf{Manual Methods:} Interviews, Questionnaires, Observation.
    \item \textbf{Automated Tools:}
        \begin{itemize}
            \item \textbf{OMR:} Reads pencil marks. \textit{(MCQ Sheets)}.
            \item \textbf{OCR:} Converts image text to editable text.
            \item \textbf{MICR:} Reads magnetic ink. \textit{(Bank Cheques)}.
            \item \textbf{Other:} Barcode Readers, RFID, Sensors.
        \end{itemize}
    \item \textbf{Input Modes:}
        \begin{itemize}
            \item \textbf{Online:} Data entered as transaction happens.
            \item \textbf{Offline:} Data collected, entered later in a batch.
        \end{itemize}
\end{itemize}
\subsectionheading{Validation (Data Quality Check)}
\begin{itemize}
    \item \textbf{Range Check:} Is value within limits? (e.g., 0-100).
    \item \textbf{Presence Check:} Is required field empty?
    \item \textbf{Type Check:} Is it the correct data type (number/text)?
\end{itemize}
\subsectionheading{Processing, Output \& Storage}
\begin{itemize}
    \item \textbf{Processing Modes:}
        \begin{itemize}
            \item \textbf{Batch:} Collect all data, process in one go. \textit{(e.g., Payroll)}.
            \item \textbf{Real-time:} Instant processing. \textit{(e.g., ATM transaction)}.
        \end{itemize}
    \item \textbf{Output Methods:}
        \begin{itemize}
            \item Direct presentation (Monitor, Printer).
            \item Storing for further processing.
        \end{itemize}
    \item \textbf{Storage Methods:}
        \begin{itemize}
            \item \textbf{Local/Remote:} On-site vs. Cloud.
            \item \textbf{Short/Long-term:} RAM vs. Hard Drive.
        \end{itemize}
\end{itemize}

\vspace{1em}
\sectionheading{4. ICT's Role \& Impact on Society}
\vspace{0.5em}
\subsectionheading{Applications of ICT}
\begin{itemize}
    \item \textbf{Education:} E-learning, CBE (Computer-Based Ed.).
    \item \textbf{Healthcare:} Diagnostics (ECG, CT), Patient records.
    \item \textbf{Agriculture:} Greenhouse control, RFID for livestock.
    \item \textbf{Business/Finance:} Banking (ATM), Payroll, Stock Mgt.
    \item \textbf{Engineering:} CAD (Design), CAM (Manufacturing).
    \item \textbf{Media:} Digital content creation, Global news access.
    \item \textbf{Law Enforcement:} CCTV analysis, GPS tracking.
\end{itemize}

\subsectionheading{Societal Impact}
\begin{itemize}
    \item \textbf{Benefits:} Economic growth, Global connectivity, Improved efficiency, Access to information.
    \item \textbf{Issues:}
        \begin{itemize}
            \item \textbf{Digital Divide:} "Haves" vs. "Have-nots" of tech access.
            \item \textbf{E-waste:} Discarded electronics. Solution: \textbf{Green Computing} (eco-friendly ICT use \& disposal).
        \end{itemize}
\end{itemize}

\subsectionheading{Ethical \& Legal Issues (The "Don'ts")}
\begin{itemize}
    \item \textbf{Confidentiality:} Keep data secret. Use passwords, encryption.
    \item \textbf{Privacy:} Right to control one's personal information.
    \item \textbf{Intellectual Property:} Legal ownership of creations.
        \begin{itemize}
            \item \textbf{Copyright:} Protects original works (books, music, code).
            \item \textbf{Piracy:} Illegal copying/distribution of copyrighted material.
            \item \textbf{Plagiarism:} Using someone's work as your own.
        \end{itemize}
    \item \textbf{Security Threats:}
        \begin{itemize}
            \item \textbf{Phishing:} Fake login pages to steal credentials.
        \end{itemize}
    \item \textbf{Netiquette:} "Network Etiquette". Good manners online.
\end{itemize}

\end{multicols}
\end{document}
